Contention management for Cloud RAN over an optical ring

### 2018-12-21 v3/articleV3.tex
version quazi finale

\graphicspath{{img/}}
\title{Contention management for Cloud RAN over an optical ring}

@@

In this article, we study a Cloud RAN (C-RAN) application based on the N-GREEN optical ring described in~\cite{ngreenarchitecture,uscumlicscalable}. C-RAN is one of the major area of development for 5G; it consists in centralizing or partially centralizing the computation units or {\bf BaseBand Units} (BBU) of the {\bf Remote Radio Heads} (RRH) in one datacenter~\cite{mobile2011c}. Periodically, each RRH on the field send some uplink traffic to its associated BBU in the datacenter, then, after a computation, the BBU send the same quantity of downlink traffic back to the RRH. The latency of the messages between the BBU and the RRH is critical since some services need end-to-end latency as low as $1$ms~\cite{3gpp5g,boccardi2014five}.

Nowadays, the traffic is managed by statistical multiplexing \todo{ref}. Here, we propose an SDN approach to {\bf deterministically} manage the periodic C-RAN traffic by choosing emission timing. Indeed, Deterministic Networking is one of the strong current aspects considered to reduce the E2E latency~\cite{finn-detnet-architecture-08}. In a previous work~\cite{dominique2018deterministic}, the authors have studied a similar problem for a star shaped network. In contrast with our previous work, finding emission timings so that different periodic sources do not use the same resource is easy in the context of the N-GREEN optical ring. However, we add two new difficulties arising from practice: the messages from RRHs are scattered and there are other traffics whose latency must be preserved. It turns out that the deterministic management of CRAN traffic
we propose reduce the latency of CRAN traffic to the physical delay of the routes, while reducing the latency of the other traffics by smoothing the load of the ring in the period. However, to achieve such a good latency, our solution needs to reserve ressource in advance, which slightly decreases the maximal load the N-GREEN optical ring can manage. Such an approach of reservation of the network for an application (CRAN in our context) relates to network slicing~\cite{jiang2016network}.

@@

        \begin{center}
      \hspace{-2.3cm}\includegraphics[scale=0.7]{containers}

   \vspace{0.5cm}

      \includegraphics[scale=0.7]{containers2}
      \end{center}

@@

   The {\bf latency} of a data packet is defined as the time it waits in an insertion buffer.
   Indeed, the routing is fixed, thus we cannot reduce the physical transmission delay of a data which depends only on the size of the arcs used. Moreover, there is only one buffering point in the N-GREEN optical ring, the insertion buffer of the node at which the data arrives. Hence, in this context, to minimize the time data need to go from source to destination, we need to minimize their (logical) latency.
   In particular, we want to reduce the latency of the C-RAN traffic to \textbf{zero}, both for the RRHs (uplink) and the BBUs (downlink). In Sec.~\ref{sec:deterministicalgorithms} we propose a deterministic mechanism with zero latency for C-RAN which also improves the latency of other data going through the optical ring. We shortly describe the nature of this additional traffic in the next subsection.

@@

The optical ring supports other traffics, corresponding to the internet flow. We call this traffic \textbf{Best Effort} (BE). We want it to have the best possible distribution of latency, but since BE traffic is less critical than C-RAN traffic, we impose no hard constraint on its latency. At each node of the ring, a {\bf contention buffer} is filled by a batch arrival process of BE data. 
\todo{dire plus précisément ce qu'est ce processus d'arrivée -> un truc qui simule du traffic de réseau avec burst en citant youssef pour les détails. Insister sur le fait que c'est donc un traffic aléatoire contrairement au C-RAN}
Then, according to the fill rate of the contention buffer and the maximum waiting time of the data, 
a packet of size at most $C$ is created by aggregating data in the contention buffer. This packet is then put in the insertion buffer of the node. Hence, the arrival of BE messages is modeled by a temporal law that gives the distribution of times between two arrivals of a packet of BE messages in a node. The computation of this distribution for the parameters of the contention buffer used in the N-GREEN optical ring is described in ~\cite{youssef2018}\todo{mettre la bonne ref}.

@@

  We first study the latency of the C-RAN and BE traffics when the ring follows an opportunistic insertion policy: When a node has a free container available, it fills it with a packet of its insertion buffer, if it exists one.
 Two different methods to manage the insertion buffer are experimentally compared. First, the \textbf{FIFO} rule, that consist in managing the C-RAN and BE packets in the same insertion buffer. Then, when a free container is available, the node fills it with the oldest packet of the insertion buffer, without distinction between C-RAN and BE. This method is compared to a method called \textbf{C-RAN prio} that uses two insertion buffers: one for the BE packets, and another for the C-RAN packets. The C-RAN insertion buffer has the priority and is used to fill containers on the ring while it is non empty before considering the BE insertion buffer.  Fig.~\ref{fig:resultopport} gives the cumulative distribution of both C-RAN and BE traffics latencies for the FIFO and the C-RAN prio methods. In this experiment, the offsets of the RRH have been randomly distributed in the period. The experimental parameters are given by Fig.~\ref{fig:params} and chosen following~\cite{ngreenarchitecture}. The results are computed over $100$ experiments (with different random offsets) where the optical ring is simulated during $10,000,000$ units of time. The source code in C of the experiments can be found on one of the authors webpage~\cite{webpage}. \todo{revoir 100 1000000}

   \vspace{0.5cm}
  \hspace{-0.75cm}

@@

\begin{minipage}[t]{.46\linewidth}

@@

\end{minipage} % ne pas sauter de ligne
\begin{minipage}[t]{.46\linewidth}
   \scalebox{.55}
  {
  \begin{tabular}{|c|c|}
  \hline
  Time to go through the cycle $RS$ & $100$ UoT \tabularnewline
  \hline
  Emission time $ET$ & $500$ UoT \tabularnewline
  \hline
   Period $P$ & $1,000$ UoT \tabularnewline
  \hline
  Number of RRH & $5$  \tabularnewline
  \hline
  Number of nodes $n$ & $5$  \tabularnewline
  \hline
   Load induced by C-RAN traffic & $50\%$  \tabularnewline
  \hline
    Load induced by BE traffic & $40\%$  \tabularnewline
  \hline
  \end{tabular}
}
\end{minipage}

  \captionof{table}{Parameters of the N-GREEN architecture}\label{fig:params}

@@

Given a ring, a set of RRH's, a period and an acceleration factor $F$, the problem we solve here is to find an \textbf{assignment} of values of the $m_i$'s which is \textbf{valid}: two RRHs never reserve or use the same container in a period. Moreover we want to preserve the latency of the BE traffic. It means that the time a BE packet spent in the insertion buffer must be minimized. To do so, we must minimize the time a node waits for a free container at any point in the period, by spreading the C-RAN traffic as uniformly as possible over the period in order to give the nodes an available container in a minimal average time.
\todo{dessin represnetation d'un assignement}

Remark that two RRHs which are not at the same position never use the same containers. Moreover, if we fix the offsets of the RRHs to even positions so that they do not reserve the same containers, then it will fix the offsets of the BBUs to odd positions which do not reserve the same containers. Hence, we need to deal with the RRHs only.

@@

Remark that the loss of containers from reservation is modest: only $RS$ by period. 
This comes from the choice of emission times of the RRHs which are taken in the order of the cycle so that there are no free containers at position zero except at the end of the period.
From this proposition, it is easy to derive the maximal number of antennas which can be supported by an optical ring, when using reservation and the same position for an RRH for the whole period.

@@

There are two sources of inefficiency in this method. The first comes from the reservation and cannot be avoided to guarantee the latency of the C-RAN traffic. The second however comes from the fact that we chose
to emit all data coming from an antenna at the same position (to guarantee zero latency). If we relax this constraint and allows to change the position of data coming from the same source during the period, 
we can use \emph{all} containers at some position and we could support $\frac{P- RS}{ET} \times \frac{F}{2}$ antennas (losing the floor in the formula).

@@

\paragraph{Packing}

For each position which is used by some RRH, and during each period, $RS$ containers are reserved before starting to emit some C-RAN traffic, while they are free. Hence, it decreases the maximal load the system can handle.
Therefore to not waste bandwidth, it is relevant to put as many RRHs as possible on the same position as in Fig.~\ref{fig:packing}. Indeed, for any position which is not used at all, $RS$ containers need not to be reserved. This strategy is also good for the latency of the BE traffic. If a position is unused, then there is always a container free of C-RAN traffic each $F$ unit of times. 
\todo{c bon deja pour la bandwith et aussi on s'assure d'avoir un container tout les F unit+ packing bof comme nom}
\begin{figure}[h!]
\begin{center}

      \includegraphics[scale=0.5]{repart0}
     \caption{Packing.}\label{fig:packing}

@@

\paragraph{Balancing inside the slot}

The free positions can be distributed uniformly over a slot, to minimize the time to wait before a node 
has access to a free container, as shown in Fig.~\ref{fig:slotbal}. It is a small optimization, since 
it decreases the latency of at most $F/2$.
\todo{dire trois mots sur les restes des restes etc}
\begin{figure}[h!]
\begin{center}

      \includegraphics[scale=0.55]{repart1}
     \caption{Balancing inside the slot.}\label{fig:slotbal}

@@

The performance when we do no packing is really bad, even worst than the methods using an insertion buffer. It is explained by the parameters of the N-GREEN ring: all positions are used and thus, there are no free containers available to BE traffic during $ET+RS$ units of time. When packing the RRHs on a minimal number of positions, the latency decreases dramatically and becomes much better than in the previous methods. The optimizations of spacing inside the slot brings almost no benefit as expected. The spacing over a period improves the latency marginally. In settings with an higher load, spacing inside a period would have a much larger effect.

In Fig.~\ref{fig:optimres}, we compare the cumulative distribution of the latency of the BE traffic using the FIFO rule or the reservation algorithm proposed here. To keep the load at $90\%$ as in the experiment of Fig.~\ref{fig:resultopport}, we set $ET = 200$ and $n = 12$. This is not out of context since the exact split of the C-RAN (the degree of centralization of the computation units in the cloud) is not fully determined yet~\cite{mobile2011c}. With these parameters, the loss of bandwidth due to reservation is at most $6\%$

@@

\vspace{0.5cm}

      \includegraphics[scale=0.25]{optim.pdf}
     \captionof{figure}{Impact of the deterministic method on the traffics.}   \label{fig:optimres}

>>>>

\graphicspath{{img/}}
\title{Contention management for Cloud RAN over an optical ring}
\makeatletter
\renewcommand*{\thetable}{\arabic{table}}
\renewcommand*{\thefigure}{\arabic{figure}}
\let\c@table\c@figure
\makeatother

@@

In this article, we study a Cloud RAN (C-RAN) application based on the N-GREEN optical ring described in~\cite{ngreenarchitecture,uscumlic2018scalable}. C-RAN is one of the major area of development for 5G; it consists in centralizing or partially centralizing the computation units or {\bf BaseBand Units} (BBU) of the {\bf Remote Radio Heads} (RRH) in one datacenter~\cite{mobile2011c}. Periodically, each RRH on the field send some uplink traffic to its associated BBU in the datacenter, then, after a computation, the BBU send the same quantity of downlink traffic back to the RRH. The latency of the messages between the BBU and the RRH is critical since some services need end-to-end latency as low as $1$ms~\cite{3gpp5g,boccardi2014five}.

Nowadays, the traffic is managed by statistical multiplexing~\cite{kern2006applying}. Here, we propose an SDN approach to {\bf deterministically} manage the periodic C-RAN traffic by choosing emission timing. Indeed, Deterministic Networking is one of the strong current aspects considered to reduce the E2E latency~\cite{finn-detnet-architecture-08}. In a previous work~\cite{dominique2018deterministic}, the authors have studied a similar problem for a star shaped network. In contrast with our previous work, finding emission timings so that different periodic sources do not use the same resource is easy in the context of the N-GREEN optical ring. However, we add two new difficulties arising from practice: the messages from RRHs are scattered and there are other traffics whose latency must be preserved. It turns out that the deterministic management of CRAN traffic
we propose reduce the latency of CRAN traffic to the physical delay of the routes, while reducing the latency of the other traffics by smoothing the load of the ring in the period. However, to achieve such a good latency, our solution needs to reserve ressource in advance, which slightly decreases the maximal load the N-GREEN optical ring can manage. Such an approach of reservation of the network for an application (CRAN in our context) relates to network slicing~\cite{jiang2016network} or virtual-circuit-switched  connections in optical networks~\cite{cadere2010virtual,szymanski2016ultra}.

@@

        \begin{center}
      \hspace{-2.3cm}\includegraphics[scale=0.65]{containers}

  % \vspace{0.5cm}

      %\includegraphics[scale=0.7]{containers2}
      \end{center}

@@

   The {\bf latency} of a data packet is defined as the time it waits in an insertion buffer.
   Indeed, because of the ring topology, the routes between RRHs and BBUs are fixed, thus we cannot reduce the physical transmission delay of a data which depends only on the size of the arcs used. Moreover, there is only one buffering point in the N-GREEN optical ring, the insertion buffer of the node at which the data arrives. Hence, in this context, to minimize the end-to-end delay, we need to minimize the (logical) latency.
   In particular, we want to reduce the latency of the C-RAN traffic to \textbf{zero}, both for the RRHs (uplink) and the BBUs (downlink). In Sec.~\ref{sec:deterministicalgorithms} we propose a deterministic mechanism with zero latency for C-RAN which also improves the latency of other data going through the optical ring. We shortly describe the nature of this additional traffic in the next subsection.

@@

The optical ring supports other traffics, corresponding to the internet flow. We call this traffic \textbf{Best Effort} (BE). We want it to have the best possible distribution of latency, but since BE traffic is less critical than C-RAN traffic, we impose no hard constraint on its latency. At each node of the ring, a {\bf contention buffer} is filled by a batch arrival process of BE data. 
This batch arrival process consist in generating, at each unit of time, a quantity of data drawn from a bimodal distribution to modelize the fact that internet traffic is sometimes bursty. Then, according to the fill rate of the contention buffer and the maximum waiting time of the data, 
a packet of size at most $C$ is created by aggregating data in the contention buffer. This packet is then put in the insertion buffer of the node. Hence, the arrival of BE messages is modeled by a temporal law that gives the distribution of times between two arrivals of a BE packet in the insertion buffer. The computation of this distribution for the parameters of the contention buffer used in the N-GREEN optical ring is described in~\cite{Cast1810:Performance}.

@@

  We first study the latency of the C-RAN and BE traffics when the ring follows an opportunistic insertion policy: When a node has a free container available, it fills it with a packet of its insertion buffer, if it exists one.
 Two different methods to manage the insertion buffer are experimentally compared. First, the \textbf{FIFO} rule, that consist in managing the C-RAN and BE packets in the same insertion buffer. Then, when a free container is available, the node fills it with the oldest packet of the insertion buffer, without distinction between C-RAN and BE. This method is compared to a method called \textbf{C-RAN priority} that uses two insertion buffers: one for the BE packets, and another for the C-RAN packets. The C-RAN insertion buffer has the priority and is used to fill containers on the ring while it is non empty before considering the BE insertion buffer.

We compare experimentally these two methods in the simplest topology: The lengths of the arcs between nodes are equal and there is one RRH by node. The experimental parameters are given in Table~\ref{fig:params} and chosen following~\cite{ngreenarchitecture}. In this experiment, the offsets of the RRH have been randomly distributed in the period. The results are computed over $1,000$ experiments (with different random offsets) where the optical ring is simulated during $1,000,000$ units of time. Fig.~\ref{fig:resultopport} gives the cumulative distribution of both C-RAN and BE traffics latencies for the FIFO and the C-RAN priority methods. The source code in C of the experiments can be found on one of the authors webpage~\cite{webpage}.
   \vspace{0.5cm}
  \hspace{-0.75cm}

@@

\begin{figure}[h]
\begin{minipage}[t]{.46\linewidth}

@@

\end{minipage} % ne pas sauter de ligne
\begin{minipage}[t]{.46\linewidth}
   \scalebox{.55}
  {
  \begin{tabular}{|c|c|}
  \hline
  Time to go through the cycle $RS$ & $100$ UoT \tabularnewline
  \hline
  Emission time $ET$ & $500$ UoT \tabularnewline
  \hline
   Period $P$ & $1,000$ UoT \tabularnewline
  \hline
  Number of RRH & $5$  \tabularnewline
  \hline
  Number of nodes $n$ & $5$  \tabularnewline
  \hline
   Load induced by C-RAN traffic & $50\%$  \tabularnewline
  \hline
    Load induced by BE traffic & $40\%$  \tabularnewline
  \hline
  \end{tabular}
}
\end{minipage}
  \captionof{table}{Parameters of the N-GREEN architecture}\label{fig:params}
\end{figure}

@@

Given a ring, a set of RRH's, a period and an acceleration factor $F$, the problem we solve here is to find an \textbf{assignment} of values of the $m_i$'s which is \textbf{valid}: two RRHs never reserve or use the same container in a period. Moreover we want to preserve the latency of the BE traffic. It means that the time a BE packet spent in the insertion buffer must be minimized. To do so, we must minimize the time a node waits for a free container at any point in the period, by spreading the C-RAN traffic as uniformly as possible over the period in order to give the nodes an available container in a minimal average time.

Figure~\ref{fig:assignment} represent an assignment of three couples RRH-BBU. each slot has a duration $F$ unit of times, and, since an RRH/BBU emits a packet each $F$ UoT during $ET$ UoT, if we take the granularity of a slot to represent the time, the emission of a BBU/RRH is continuous in the time, during $ET$ UoT. Also, if we take a date $t$ in the period $P$, then $t$ is in the slot $\frac{P}{F}$ and is at position $P \mod F$.

      \includegraphics[scale=0.65]{assignment}
     \caption{Representation of an assignment with $F = 6$.}\label{fig:assignment}

Remark that two RRHs which are not at the same position never use the same containers. Moreover, if we fix the offsets of the RRHs to even positions so that they do not reserve the same containers, then, because the answers of the BBU are sent without delay in our model, it will fix the offsets of the BBUs to odd positions which do not reserve the same containers. Hence, we need to deal with the RRHs only.

@@

Remark reserving free containers make them unusable for BE traffic which is akin to a loss of bandwidth. However, with our choice of emission times of the RRHs in the order of the cycle, we reserve $ (k-1)\times RS$ containers used by other RRHs and only $RS$ free containers are wasted.

From Proposition~\ref{prop:assign}, it is easy to derive the maximal number of antennas which can be supported by an optical ring, when using reservation and the same position for an RRH for the whole period.

@@

There are two sources of inefficiency in this method. The first comes from the reservation and cannot be avoided to guarantee the latency of the C-RAN traffic. The second however comes from the fact that an RRH must emit at the same position during all the emission time (to guarantee zero latency). If we allow to change the position of data coming from the same RRH during the period, 
we can use \emph{all} containers in a position and hence we could support $\lfloor\frac{P- RS}{ET} \times \frac{F}{2}\rfloor$ antennas.

@@

\paragraph{Compacting positions}

For each position which is used by some RRH, and during each period, $RS$ containers are reserved before starting to emit some C-RAN traffic, while they can be free. Hence, it decreases the maximal load the system can handle.
Therefore to not waste bandwidth, it is relevant to put as many RRHs as possible on the same position as in Fig.~\ref{fig:packing}. Indeed, for any position which is not used at all, $RS$ containers need not to be reserved. This strategy is also good for the latency of the BE traffic. If a position is unused, then there is always a container free of C-RAN traffic each $F$ unit of times. 
\begin{figure}[h!]
\begin{center}

      \includegraphics[scale=0.5]{repart0}
     \caption{Compacting positions.}\label{fig:packing}

@@

\paragraph{Balancing used positions}

The free positions can be distributed uniformly over a slot, to minimize the time to wait before a node 
has access to a free container, as shown in Fig.~\ref{fig:slotbal}. To do so, compute the number of needed frequencies $x = \lceil k\times \frac{ET}{P - RS}\rceil$, with k the number of antennas. Then, choose the $x$ positions in the following way: $\frac{F}{x} -1 $ free positions are set between each used positions. If $\frac{F}{x}$ has a reminder, then let one more free position between the $F \mod x$ first used position. It is a small optimization, since 
it decreases the latency of at most $F/2$.

      \includegraphics[scale=0.55]{repart1}
     \caption{Balancing used positions.}\label{fig:slotbal}

@@

The performance when we do no compacting positions is really bad, even worst than the methods using an insertion buffer. It is explained by the parameters of the N-GREEN ring: all positions are used and thus, there are no free containers available to BE traffic during $ET+RS$ units of time. When compacting the RRHs on a minimal number of positions, the latency decreases dramatically and becomes much better than in the previous methods. The optimizations of spacing inside the slot brings almost no benefit as expected. The spacing over a period improves the latency marginally. In settings with an higher load, spacing inside a period would have a much larger effect.

In Fig.~\ref{fig:optimres}, we compare the cumulative distribution of the latency of the BE traffic using the FIFO rule or the reservation algorithm proposed here. To keep the load around $90\%$ as in the experiment of Fig.~\ref{fig:resultopport}, we set $ET = 200$ and $n = 12$. This is not out of context since the exact split of the C-RAN (the degree of centralization of the computation units in the cloud) is not fully determined yet~\cite{mobile2011c}. With these parameters, the loss of bandwidth due to reservation is at most $6\%$

@@

\vspace{0.5cm}
\todo{packing seul programmé, a voir comment on l'integre dans les resultats}
      \includegraphics[scale=0.25]{optim.pdf}
     \captionof{figure}{Impact of the deterministic method on the traffics.}   \label{fig:optimres}
